Exploring probability distributions for bivariate temporal granularities

### 2019-12-15 main.Rmd
interchangeability section update

## Levels within facet and aesthetic variables

@@

## Interaction of facet and aesthetic variables

@@

## Statistical transformation computation

Even with harmonies, visualizing probability distributions can be misleading either due to rarely occuring categories or unevenly distributed events.

@@

Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories. Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times. There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$. If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable. The value of $k$ will depend on what type of visualization we are producing. For a decile plot, even $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$. Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.

@@

Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel. This might happen due to missing observations in the data or uneven locations of events in time domain. In such cases, the visualizations that use density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators. \autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.

>>>>

## Facet and aesthetic variables

### Levels

@@

### Interaction

@@

### Interchangeability

When we consider the layers of grammars considered for our data structure <$A_i$, $B_j$, $v(s)$> $i \in {1, 2, \dots, n}$ and $j \in {1, 2, \dots, m}$, mapping $A_i$ mapped to $x$ position and $B_j$ to facets
suggests that levels of $A_i$'s which are placed in close proximity are to be compared against each other and each level of $B_j$ represent a group. When $A_i$ mapped to facets and $B_j$ is mapped to $x$ axis, distributions are compared across levels of $B_j$ and levels if $A_i$ start acting as a group. However, the variables used in the layers are the same, the difference in mapping emphasizes different behavior of the variables.

## Number of observations and statistical transformations

Even with harmonies, visualizing distributions can be misleading if statistical transformations are performed on rarely occuring categories or unevenly distributed events.

@@

Suppose we have $T$ observations, and two cyclic granularities $C_1$ with $n$ categories and $C_2$ with $m$ categories. Each element of $C_1$ occurs approximately $T/n$ times while each element of $C_2$ occurs approximately $T/m$ times. There are no empty combinations, and each combination will occur on average an equal number of times as $T\rightarrow\infty$, so the average number of observations per combination is $T/(mn)$. If we require at least $k$ observations to create a meaningful panel, then provided $T \ge mnk$, the visualization will be acceptable. The value of $k$ will depend on the statistical transformation we are producing. For computing deciles, $k=10$ may be acceptable, but for density estimates, we would need $k \ge 30$. Rarely occurring categories such as the 366th day of the year, or the 31st day of the month can suffer from such problem.

@@

Even when there are no rarely occuring events, number of observations might vary hugely within or across each facet. This might happen due to missing observations in the data or uneven locations of events in time domain. In such cases, the statistical transformations based on density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators. \autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.
